Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2018-10-16 tex/WW4C_VHzQ.tex
tester.bib added

In this paper we include all non-proprietary WFCAM data, which covers all
public surveys and PI projects from Semester 05A to 1st January 2017 and all
non-proprietary VISTA data, which covers all public surveys and PI projects from
science verification (20091015) to 1st April 2016, to get as much coverage as
possible. Hence, the NIR coverage is extremely heterogeneous.

The data was processed using a forced photometry, or matched-aperture
photometry. A pipeline for this has been developed by WFAU, but
has not yet been incorporated into the main VDFS pipeline. Full details of the
matched-aperture pipeline (MAP) will appear in a forthcoming paper, Cross et al.
2018, in prep, and has also been discussed in \citet{Cross2013}.  
%% and \citet{Cross2017} -- Can't find Cross et al. 2017 in NASA/ADS. [NJC: this was a conference proceeding
%% from ADASS2016. They just contacted me about the copyright. They are so slow even though we 
%% had to submit our proceedings before the conference. I had assumed it was already published.]
We will give the pertinent points of the MAP and specific details of these data sets below.

The first step of the process was to ingest the catalogue of VHzQs into
the WSA/VSA. These data are stored in the table \verb+finalQsoCatalogue+. Next we set
up new programmes WSERV1000, VSERV1000 in the WSA and VSA respectively. We
assign all frames that potentially contain the high-z QSO based on WCS information in the
WSA to WSERV1000, but adding a new line to the table \verb+ProgrammeFrame+ and
same in VSA for VSERV1000. Assigning these frames to a separate programme allows
us to process them together and put them into a single release.

\subsubsection{Setup Matched Aperture Product}

To create matched aperture products it is first necessary to define the product. We do this in two requirements tables, \verb+RequiredMatchedApertureProduct+ and \verb+RequiredMapAverages+ that
tell the \textbf{ProgrammeBuilder} software which tables to create and what type of product is being created. These requirements tables also drive the pipeline.

\verb+RequiredMatchedApertureProduct+ contains the following attributes for the High-z QSO MAP product.
\begin{itemize}
    \item mapID, Identifier of the MAP product, value=1	
    \item archiveName, Main archive of product, value='VSA'/'WSA' 	
    \item programmeID, UID of programme, in this case a combined programme, value=10999
    \item extractor, Extractor used for forced photometry, value='CASU' (imcore_list)
    \item mapType, Type of MAP product (options 0 internal band-merged, 1 band-merged across surveys, 2 multi-epoch), value=2 
    \item name, Name of MAP product, value='highzQsoMap'
    \item description, Description of MAP product, 'Variability via matched apertures for set of high-z QSOs'	
    \item selection, Selection, e.g. File, pseudo-SQL, FITS image that is used to give input list, value='SELECTSTR qsoID,ra,dec FROMSTR finalQsoCatalogue'	
    \item finalProductTable, Name of final product table, value='highzQsoMapVariability'
    \item addSurveyList, List of additional archive/surveys that data comes from, value='NONE'
\end{itemize}

\verb+RequiredMatchedApertureProduct+ gives all of the information such as the {\it programmeID} (identifying which data), the {\it mapType} (type of processing), the {\it extractor} to use, the input list {\it selection}, which in this case is the IDs and positions of the QSOs in \verb+finalQsoCatalogue+ which contains all the literature data and pre-derived parameters for the 424 QSOs.

\verb+RequiredMapAverages+ is used to setup the averaging of the MAP photometry from several images. We can average the epochs in multiple ways: over all the input images to a deep image, or similarly the input pawprint images to tile image in VISTA. Alternatively we can average over a fixed number of epochs or a fixed duration in time. Multiple setups can be defined in \verb+RequiredMapAverages+, and the different averages selected using the defined {\it setupID}.

    Add table?
    \begin{table*}
      \begin{center}
        \setlength{\tabcolsep}{4pt}
        \begin{tabular}{ll lll lll l}
          \hline \hline
   programmeID &	mapID &	setupID  &	description               &	useDeeps  &	useHighProd &	timeScale	    &  nEpochs     &	overLaps\\
          \hline
10999	& 1 & 0 & Average over whole time & 0 &	NONE &	-1.000000 &	-99999999 &	1 \\
10999	& 1 & 1 & Average over a week &	0 & NONE & +7.000000 &	-99999999 &	1\\
10999	& 1 & 2	& Average over a fortnight & 0 & NONE &	+14.000000 & -99999999 & 1\\
10999	& 1 & 3	& Average over a month & 0 & NONE & +30.000000 & -99999999 & 1\\
10999	& 1 & 4	& Average over 10 epochs & 0 & NONE & -9.999995E008 & 10 & 1\\
10999	& 1 & 5 & Average over 6 months	& 0 & NONE & +183.000000 & -99999999 & 1 \\
10999	& 1 & 6	& Average over 1 year &	0 & NONE & +365.000000 & -99999999 & 1 \\
10999	& 1 & 7	& Average over 2 years & 0 & NONE & +730.000000	& -99999999 & 1 \\
10999	& 1 & 8 & Average over 3 months & 0 & NONE & +91.000000	& -99999999 & 1 \\
          \hline \hline
          \label{tab:reqMapAvg}
        \end{tabular}
        \caption{}
      \end{center}
    \end{table*}

    \subsubsection{Processing MAP data and Forced photometry} 
    A set of apertures to do forced photometry with were created in 
    \verb+MapApertureIDshighzQsoMap+, based on the \verb+RequiredMatchedApertureProduct+ 
    {\it selection} value, which in this case is to use the known positions of $z>5$ QSOs. This may
    seem unnecessary layer since it is a one-to-one match with \verb+finalQsoCatalogue+,
    but the layer is a general abstraction that allows multiple selections of data, including complex queries
    or input images.

    CASU's \verb+imcore_list+ was run on each frame assigned to the programme with
    the input list equal to objects selected to be within \ldots of the multiframe
    centre. In this case, where there are 424 QSOs spread across the sky, there is
    typically 1 object per multiframe, but many have two and one frame,
    multiframeID=2614475 in the VSA (a J-band stack from a P90 PI programme), has 3 objects. 
    The output of \verb+imcore_list+
    is a multi-extension FITS binary table, with identical columns to the CASU FITS
    binary tables produced using the CASU \verb+imcore+\footnote{http://casu.ast.cam.ac.uk/surveys-projects/vista/technical/catalogue-generation}, the basis of all WFCAM and
    VISTA survey catalogue products and also familiar to many astronomers through the INTWFS, 
    VST-ATLAS and VPHAS. Since forced photometry can be run
    in many ways on a particular image depending on the input catalogue, a single science frame is not
    necessarily linked to a single MAP catalogue, so the \verb+MapFrameStatus+ table
    links images ({\bf multiframeID}) to catalogues ({\bf catalogueID}).
    For matched aperture measurements, we process these binary tables using the same
    software, as the standard extracted catalogues, but implement a few different
    processes. We calculate both Luptitudes \citep{Luptitudes}, and calibrated fluxes,
    see Eqn~\ref{eq:jky}.

    \begin{equation}
      F^{calib} = f\times10^{-0.4*(ZP+VAB-8.90+corr)} 
      \label{eq:jky}
    \end{equation}

    \noindent where $f$ is the measured flux in ADU, $ZP$ is the zeropoint of
    frame, $VAB$ is the Vega-to-AB correction for the filter, and $corr$ are the
    various correction terms, e.g. distortion correction, scattered light
    correction, aperture correction, see \citep{Hambly2008, Cross2012}.

    Once all data is processed, the data are ingested into
    \verb+wserv1000MapRemeasurement+ and \verb+vserv1000MapRemeasurement+
    respectively. These tables are similar in structure to \verb+lasDetection+
    described in \citep{Hambly2008} or other UKIDSS or VISTA detection tables. One of
    the main differences is the primary key which is ({\bf apertureID},{\bf
      catalogueID}) rather than ({\bf multiframeID},{\bf extNum},{\bf seqNum}),
    reflecting the fact that these are forced photometry on a known set of apertures
    and a set of catalogues, rather than detections above a threshold in a set of
    detector extensions to image multiframes.

    Quasars are known to vary (both photometrically and spectroscopically)
and very high-$z$ quasars under going super-Eddington accretion during
a rapid BH growth phase are prime candidates for this variation. [NJC: This is better placed somewhere else. We first used it to justify averaging over ~4 weeks, but then decided to average over many periods and select later.]
Therefore, with repeat observations over many epochs available via the
WSA, we have a choice to make for how we report the photometry. The remeasured photometry is averaged to
improve signal-to-noise over all the averaging schemes setup in Table~\ref{tab:reqMapAvg}. This allows us to get the best combination of signal-to-noise and number of epochs for different QSOs.

    \subsubsection{Averaging matched photometry}
    Unfortunately the photometry in a single epoch image often has low
    signal-to-noise. The advantage of matched aperture photometry on QSOs is that
    co-adding is relatively simple: if each epoch is taken in the same aperture
    (assuming the astrometry is correct) and the aperture photometry has been
    corrected to total: the standard aperture corrections work well for point
    sources then averaging the calibrated photometry should give as good as
    extracting from a deep stack. In fact it will almost certainly do better, since
    the images in some cases are taken from multiple projects with different
    pointings, and orientations and exposures, so stacking may not be optimal. 
    In addition, corrections such as scattered light, pixel distortion
    and aperture corrections will be more complex, if not impossible to apply to
    deep-stacks but will be automatically applied in the averaging of forced
    photometry.

    We average the aperture corrected calibrated fluxes (e.g. {\bf aperJky3}), and
    then convert to magnitudes or luptitudes. Since we do not have a deep image for
    each set of averages, we cannot calculate non-aperture corrected values, so the
    photometry is only appropriate for point-sources.

    \begin{equation}
      \bar{F} = \frac{\sum_i^N (w_i\,F_i)}{\sum_i^N w_i}  
      \label{eq:avg}
    \end{equation}

    \noindent where $F_i$ is the $i^{th}$ epoch measurement of a parameter to be
    averaged such as the aperture corrected calibrated flux in a $1\arcsec$ aperture
    ({\bf aperJky3}) and $\bar{F}$ is the weighted mean average of this parameter.
    The weight for each epoch $w_i=1/(\sigma_{F})^2$ if the epoch is included and 
    $w_i=0$ if an epoch is excluded for quality control purposes. We exclude inputs
    where the bitwise quality flag, {\bf ppErrBits}>256, and additionally, in the
    case of VISTA, we exclude data on detector 16 ({\bf extNum=17}), where the 
    quantum efficiency is variable. If all remeasurements are excluded, then one  of
    the ones with the lowest {\bf ppErrBits} is used. In this case the averages 
    will be equal to the original measurement, and the reader may wonder why we 
    bother to duplicate the data. It is important to do this to simplify queries 
    for users who will then not have to set up complex queries to work out where 
    the data is, and only around $2\%$ of VISTA MAP measurements were excluded  so
    the fraction of duplications is very low. The average data is stored in 
    \verb+wserv1000MapRemeasAver+ for WFCAM and \verb+vserv1000MapRemeasAver+ for 
    VISTA, which again are organised with a primary key of ({\bf apertureID},{\bf catalogueID}).
    Some parameters that are averaged do not have an error $\sigma_{F}$, so in these
    cases we weight them using the {\bf averageConf} parameter, which has a median
    of 100. In the case of WFCAM where this is not calculated, we set each epoch to
    have an {\bf averageConf}$=100$. The {\bf sumWeights} parameter in the
    \verb+MapRemeasAver+ tables is the sum of all the {\bf averageConf} values that
    went into the calculation and the \verb+vserv1000MapAverageWeights+ table gives
    the weights for each epoch, linked to each average for VSERV1000.

    We calculate a set of averaged catalogues, for each pointing and filter, based
    on the requirements in \verb+RequiredMapAverages+, in these cases over time
    spans of 7, 14, 30, 91, 183 days, 365 days, 730 days, over 10 epochs and over all epochs,
    see Tab.~\ref{tab:reqMapAvg}. The
    averaging process starts at the first epoch and works on. Possible future
    refinements could first look for denser groupings that satisfied the criteria so
    that these were not split into two groups. For this high-z QSO project we have averaged over overlaps, 
    because in several cases data for one source comes from multiple original projects that have slightly 
    different pointings.

    Each averaged catalogue has a separate entry in
    \verb+MapFrameStatus+ and is linked to each original epoch remeasured catalogue
    through \verb+MapProvenance+, which contains the averaged catalogue
    identifier {\bf combiCatID}, the epoch catalogue identifier {\bf catalogueID}
    and the averaging setup ID {\bf avSetupID}.

We give our recipe and SQL query syntax in Appendix~\ref{sec:SQL}.

    The code to calculate variability statistics as described in \cite{Cross2009} has
    not yet been completed to run with the matched aperture pipeline, so these are
    not included. Variables were selected ...

@@

    \begin{figure}
      \centering
      \includegraphics[width=8.5cm]
      {/cos_pc19a_npr/programs/quasars/highest_z/MIR_LCs/NEOWISER_LC_histogramlog_20180827.png}
      \vspace{-16pt}
      \caption[]
      {Histogram showing the number of NEOWISE-R epochs and detections there are for each 
        VH$z$Q.} 
      \label{fig:MIR_LC_epochs}
    \end{figure}

@@

\begin{figure}
  \includegraphics[width=8.5cm]
  {/cos_pc19a_npr/programs/quasars/highest_z/MIR_LCs/three_MIR_LC_egs_20180827.pdf}
  \centering
  \caption[]
  {Here we show the MIR NEOWISE-R for J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751. 
    Red points are the W1 band; cyan points the W2 band.} 
  \label{fig::MIR_LC_3egs}
\end{figure}

@@

@@

@@

@@

\begin{enumerate}
\item http://wsa.roe.ac.uk/ 
\item Login
\item {\tt username:	WSERV1000;  password: 	highzqso;   community: 	nonSurvey} 
\item Freeform SQL Query with  WSERV1000v20180327
\end{enumerate}

@@

%\bibliographystyle{apj}
\bibliographystyle{mn2e}
\bibliography{/cos_pc19a_npr/LaTeX/tester_mnras}

>>>>

In this paper we include all non-proprietary WFCAM data, which covers all
public surveys and PI projects from Semester 05A to 1st January 2017, and all
non-proprietary VISTA data, which covers all public surveys and PI projects from
science verification (20091015) to 1st April 2016.

The data was processed using a matched-aperture photometry method
where flux is recovered at the spectroscopic position of the quasar,
without necessarily knowing if there is a formal detection in the NIR
photometry beforehand.  Full details of the matched-aperture pipeline
will appear in a forthcoming paper, Cross et al.  2018, in prep, and
has also been discussed in \citet{Cross2013}.

We query the WSA and VSA performing matched-aperature photometry at
the positions of our 424 VH$z$Qs. We return a database with $xxxx$
entries. This database is world-readable and we give the full receipe
and relevant SQL queries in Appendix~\ref{sec:SQL}.

    \subsubsection{Averaging matched photometry}
    The photometry in a single epoch image often has low
    signal-to-noise.  The advantage of matched aperture photometry on QSOs
    is that co-adding is relatively simple if each epoch is taken in the
    same aperture and the aperture photometry has been corrected to
    total. Indeed, the standard aperture corrections work well for point
    sources. Coadding using the matched-aperature photometry, where the
    individual epochs are taken from multiple projects with different
    pointings and orientations, should help with issues such as scattered
    light, pixel distortion and aperture corrections.

    We average the aperture corrected calibrated fluxes (e.g. {\bf aperJky3}), and
    then convert to magnitudes. Since we do not have a deep image for
    each set of averages, we cannot calculate non-aperture corrected values, so the
    photometry is only appropriate for point-sources.

    \begin{equation}
      \bar{F} = \frac{\sum_i^N (w_i\,F_i)}{\sum_i^N w_i}  
      \label{eq:avg}
    \end{equation}
    where $F_i$ is the $i^{th}$ epoch measurement of a parameter to be
    averaged such as the aperture corrected calibrated flux in a $1\arcsec$ aperture
    ({\bf aperJky3}) and $\bar{F}$ is the weighted mean average of this parameter.
    The weight for each epoch $w_i=1/(\sigma_{F})^2$ if the epoch is included and 
    $w_i=0$ if an epoch is excluded for quality control purposes.

    We calculate a set of averaged catalogues, for each pointing and filter, based
    on the requirements in \verb+RequiredMapAverages+, in these cases over time
    spans of 7, 14, 30, 91, 183 days, 365 days, 730 days, over 10 epochs and over all epochs,
    see Tab.~\ref{tab:reqMapAvg}. The
    averaging process starts at the first epoch and works on.
    %%
    We give examples of the SQL queries syntax in Appendix~\ref{sec:SQL}.

    114 quasars have both WFCAM and VISTA data.

@@

    \begin{figure}
      \centering
      \includegraphics[width=8.5cm]
      {/cos_pc19a_npr/programs/quasars/highest_z/LightCurves/MIR_LCs/NEOWISER_LC_histogramlog_20180827.png}
      \vspace{-16pt}
      \caption[]
      {Histogram showing the number of NEOWISE-R epochs and detections there are for each 
        VH$z$Q.} 
      \label{fig:MIR_LC_epochs}
    \end{figure}

@@

\begin{figure}
  \includegraphics[width=8.5cm]
  {/cos_pc19a_npr/programs/quasars/highest_z/LightCurves/MIR_LCs/three_MIR_LC_egs_20180827.pdf}
  \centering
  \caption[]
  {Here we show the MIR NEOWISE-R for J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751. 
    Red points are the W1 band; cyan points the W2 band.} 
  \label{fig::MIR_LC_3egs}
\end{figure}

@@

      $J_{\rm 2MASS}$       &   12350   &       10806  & 	14068  &  1624       &  $J_{\rm 2MASS}  $     & $= J_{AB}    - 0.894  $         \\

@@

      $H_{\rm 2MASS}$      & 16620        & 	14787  &   18231   & 2509      & $H_{\rm 2MASS}$      & $= H_{AB}  - 1.374 $        \\

@@

      $K$s$_{\rm 2MASS}$     &   21590    & 	19544  &   23552        &   2619      & $K$s$_{\rm 2MASS}$      & $ = K$s$_{AB} -  1.84  $          \\ 

@@

The data are on the WFCAM Science Archive: \href{wsa.roe.ac.uk}{\tt wsa.roe.ac.uk}. 
Access the User Login form \href{WFCAM Science Archive}{\tt wsa.roe.ac.uk/login.html} 
with these credentials::
\begin{itemize}
    \item Username: {\tt WSERV1000}
    \item password: {\tt highzqso}
    \item community: {\tt nonsurvey}
\end{itemize}
Then going to the \href{http://wsa.roe.ac.uk:8080/wsa/SQL_form.jsp}{[\tt Free Form SQL 
Query} page the Database release {\tt WSERV1000v20180716} can be accessed which 
contains all the data we use here.

@@

%\bibliographystyle{apj}
\bibliographystyle{mn2e}
\bibliography{tester_mnras}
